---
title: "PML"
output: html_document
---


##Executive Summary
6 participants were asked to perform barbell lifts correctly and incorrectly in 5 different ways. These information is used to predict the movementof 20 test cases.

A random tree method was used to predict results at an accuracy of 0.998. The 20 test rusults were predicted correctly at a rate of 19/20 (95%).

##Setup
Assuming that the csv files are in the directory:
```{r}
library(caret)
library(randomForest)
library(rpart) 
library(RColorBrewer)
library(lattice)
```

```{r}
setwd('C:/Workspace/pml')#change as appropriot 
training <- read.csv('pml-training.csv')
testing <- read.csv('pml-training.csv')
```

##Partition for cross validation
60% partition is used for training and rest for testing

```{r}
partitions <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
partTrain <- training[partitions, ]
partTest <- training[-partitions, ]
dim(partTrain); 
dim(partTest)
```

##Cleaning Data
excluded identifier, timestamp, and window data and those variables with higher than 95% missing values.
```{r,results='hide',warning=FALSE,message=FALSE}
Clean <- grep("name|timestamp|window|X", colnames(training), value=F) 
trainingCl <- partTrain[,-Clean]
trainingCl[trainingCl==""] <- NA
NA.rate <- apply(trainingCl, 2, function(x) sum(is.na(x)))/nrow(trainingCl)
trainingCl <- trainingCl[!(NA.rate>0.95)]
summary(trainingCl)
```
PCA is al=pplied to reduce the number of variables further
```{r,results='hide',warning=FALSE,message=FALSE}
preProc <- preProcess(trainingCl[,1:52],method="pca",thresh=.95) #25 components are required

preProc <- preProcess(trainingCl[,1:52],method="pca",pcaComp=25) 
preProc$rotation
trainingPC <- predict(preProc,trainingCl[,1:52])
```

##Random forest
Random forest method is applied
```{r}
modFitRF <- randomForest(trainingCl$classe ~ .,   data=trainingPC, do.trace=F)
print(modFitRF)
```

##Checking test set
```{r}
testingCl <- testing[,-Clean]
testingCl[testingCl==""] <- NA
NA.rate <- apply(testingCl, 2, function(x) sum(is.na(x)))/nrow(testingCl)
testingCl <- testingCl[!(NA.rate>0.95)]
testingPC <- predict(preProc,testingCl[,1:52])
confusionMatrix(testingCl$classe,predict(modFitRF,testingPC))
```

##Running the 20 samples
```{r,results='hide',warning=FALSE,message=FALSE}
testingCl <- testing[,-Clean]
testingCl[testingCl==""] <- NA
NA.rate <- apply(testingCl, 2, function(x) sum(is.na(x)))/nrow(testingCl)
testingCl <- testingCl[!(NA.rate>0.95)]
testingPC <- predict(preProc,testingCl[,1:52])
testingCl$classe <- predict(modFitRF,testingPC)
```

##Conclusion
From the 19622 observations 60% were used to build a random tree. A 0.9894 accuracy is achieved with the testing set. The prediction for the 20 smaple rates yieled 1 incorrect result. 
